EldersAIR initial Micropem data work

### 2020-09-24 micropem_data_cleaning.Rmd
micropem cleaning

npt_micropem_new <- npt_micropem %>% 
  group_by(mp_file_name) %>% 
  mutate(pm_personal_new = if_else(pm_personal < 0, 0, pm_personal),
         pm_mean_new = mean(pm_personal_new, na.rm = TRUE),
         pm_diff = round(pm_mean_new - pm_mean, digits = 2),
         pm_diff_perc = round(abs(pm_diff/pm_mean*100), digits = 2),
         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 2, 1, 0)) %>% 
  ungroup() %>% 
  distinct(mp_file_name, .keep_all = TRUE) %>% 
  select(mp_file_name, area, home, home_winter_id, sampling_visit,
         pm_mean, pm_mean_new, pm_diff, pm_diff_perc, pm_indicator) %>% 
  filter(pm_indicator == 1)
```

>>>>

# https://cran.r-project.org/web/packages/rtimicropem/vignettes/chai_data_cleaning.html 
# The method in the above link holds off negative values and checks change in mean
# They check to see if mean changed by >5%, and don't use the file if it did

npt_summary <- npt_micropem %>% 
  distinct(mp_file_name, .keep_all = TRUE) %>% 
  summarize(mean_pm = mean(pm_mean, na.rm = TRUE), 
            sd_pm = sd(pm_mean, na.rm = TRUE),
            min_pm = min(pm_mean, na.rm = TRUE),
            med_pm = median(pm_mean, na.rm = TRUE),
            max_pm = max(pm_mean, na.rm = TRUE))

npt_summary_new <- npt_micropem %>% 
  group_by(mp_file_name) %>% 
  mutate(pm_personal_new = if_else(pm_personal < 0, 0, pm_personal),
         pm_mean_new = mean(pm_personal_new, na.rm = TRUE),
         pm_diff = round(pm_mean_new - pm_mean, digits = 2),
         pm_diff_perc = round(abs(pm_diff/pm_mean*100), digits = 2),
         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 3, 1, 0)) %>% 
  ungroup() %>% 
  distinct(mp_file_name, .keep_all = TRUE) %>% 
  select(mp_file_name, area, home, home_winter_id, sampling_visit,
         pm_mean, pm_mean_new, pm_diff, pm_diff_perc, pm_indicator,
         pm_sample_interval) %>% 
  filter(pm_indicator == 0) %>% 
  filter(pm_sample_interval > 1.6)
  summarize(mean_pm = mean(pm_mean_new, na.rm = TRUE), 
            sd_pm = sd(pm_mean_new, na.rm = TRUE),
            min_pm = min(pm_mean_new, na.rm = TRUE),
            med_pm = median(pm_mean_new, na.rm = TRUE),
            max_pm = max(pm_mean_new, na.rm = TRUE))

npt_summary
npt_summary_new

# Above, I changed all negative PM values to 0 and checked the change in mean PM per file
# Some files had small absolute changes in mean PM, but since levels were so low,
# the percent changes were very high. I filtered out files with the following criteria:
# -After changing negative values to 0, if percent change in mean PM was >5%,
# -and if the absolute difference in mean PM was >3 ug/m3
# I also calculated overall summary stats before and after filtering out these files.
# Mean and median PM changed by less than 1 ug/m3. SD was essentially unchanged.
# By using this method, I filter out files that are highly impacted by negatives.
# The remaining files are not as impacted by negatives, and overall summary stats are
# impacted minimally.
## This method filtered out 7 files.
# I also filter out files that had a run time <80% expected.
## This method filtered out 9 additional files, for a total of 16 files removed.
# The steps below will use the methods on the full dataset

npt_micropem_new <- npt_micropem %>% 
  group_by(mp_file_name) %>% 
  mutate(pm_personal_new = if_else(pm_personal < 0, 0, pm_personal),
         pm_mean_new = mean(pm_personal_new, na.rm = TRUE),
         pm_diff = round(pm_mean_new - pm_mean, digits = 2),
         pm_diff_perc = round(abs(pm_diff/pm_mean*100), digits = 2),
         pm_indicator = if_else(pm_diff_perc > 5 & pm_diff > 3, 1, 0)) %>% 
  ungroup() %>% 
  filter(pm_indicator == 0) %>% 
  filter(pm_sample_interval > 1.6)

write_rds(npt_micropem_new, paste0(file_path, "Output/npt_micropem_new3.rds"))
```
